\TheTitle

### 2019-12-13 Paper1/results.tex
pipe testcase first draft; figs not final

When looking at the computed IPM expectation value and variance (see Figure~\ref{fig:reados2DE} and \ref{fig:reados2DVar}) and comparing the results against the reference solution in Figure~\ref{fig:referenceSolutionsub2DE} and \ref{fig:referenceSolutionsub2DVar}, one can observe that since the uncertainties have a bigger effect on the result than in the one-dimensional case, the IPM solution still shows a step-like profile. However, we are able to capture the main features of the reference solution. The increased effect of the uncertainty can again be seen in Figure~\ref{fig:readosIPMEVar2DRefinement}, where a refinement level of $8$, i.e. a truncation order of $9$ is chosen on a larger portion than for the one-dimensional case. \remja{Visualizing the variance on a logarithmic scale in Figure~\ref{fig:referenceSolutionsub2DVarLog} further shows that the applied refinement indicators work well and use higher order moments in areas of high variance.}

\subsection{Pipe Testcase}
\begin{figure}[H]
	\centering
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/pipe/pipe_sc_ref_n50_E(rho).png}
		\label{fig:referenceSolutionsPipe1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/pipe/pipe_sc_ref_n50_Var(rho).png}
		\label{fig:referenceSolutionsPipe2}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/pipe/pipe_mesh.png}
		\label{fig:referenceSolutionsPipe2}
	\end{subfigure}
	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh.}
	\label{fig:referenceSolution3D}
\end{figure}
\comment{TODO: Pipe testcase}

>>>>

When looking at the computed IPM expectation value and variance (see Figure~\ref{fig:reados2DE} and \ref{fig:reados2DVar}) and comparing the results against the reference solution in Figure~\ref{fig:referenceSolutionsub2DE} and \ref{fig:referenceSolutionsub2DVar}, one can observe that since the uncertainties have a bigger effect on the result than in the one-dimensional case, the IPM solution still shows a step-like profile. However, we are able to capture the main features of the reference solution. The increased effect of the uncertainty can again be seen in Figure~\ref{fig:readosIPMEVar2DRefinement}, where a refinement level of $8$, i.e. a truncation order of $9$ is chosen on a larger portion than for the one-dimensional case. Visualizing the variance on a logarithmic scale in Figure~\ref{fig:referenceSolutionsub2DVarLog} further shows that the applied refinement indicators work well and enforce the usage of higher order moments in areas of high variance.

\subsection{2D Euler equations with a three dimensional uncertainty}
For the last numerical study we again use the Euler equations, but this time not applied to the NACA airfoil, but rather to a bend Sod shock tube experiment. The prescribed initial condition describes a gas at rest, but with a discontinuity, i.e. as sudden jump, in density as well as energy in the upper part of the tube. The density $\rho_u = p/(R \cdot T)$ in the upper part results in $\rho_u = 1.289$ with $p = 101\;325$ Pa, a temperature of $273.15$ K and the specific gas constant for air $R=287.87$. The energy $\rho e_u$ is set to exactly $1.0$. For the initial conditions in the lower part we set $\rho_l = 0.5 \cdot \rho_u$ and $\rho e_l = 0.3$. The heat capacity ratio $\gamma$ is set to $1.4$ as in the previous studies. We again augment the deterministic case by inflicting three uncertainties this time. The applied boundary conditions are Euler slip conditions $\bm{v}^T\bm{n}=0$ for the walls of and Dirichlet conditions set to the initial conditions at each end of the tube. The first two uncertainties are the density and energy of the lower part of the shock tube. Here we set $\rho_l \sim U(1.189,1389)$ and $\rho e_l \sim U(0.2,0.4)$. The third uncertainty enters through the position of the shock itself and thus $y_\text{shock} \sim U(1.0, 1.2)$. Note, that in contrast to the upper testcases, here we are not interested in the steady state of the system but rather the evolution of the expectancy and variance in time. The computational mesh (see Figure \ref{fig:referenceSolutionsPipeMesh}) is an unstructured mesh of $25458$ triangular cells, where all the cells are similar in size. The simulations were run until a time of $0.2$s was exceeded and the corresponding results can be seen in Figure \ref{fig:solution3D}. The reference solution in Figure \ref{fig:referenceSolutionsPipe1} and \ref{fig:referenceSolutionsPipe2} was again computed using SC with 50 quadrature points in each stochastic dimension, yielding a total of $50^3=125\;000$ quadrature points.\\
For this testcase we want to compare the results for sparse grids w.r.t. tensorized grids in higher stochastical dimensions. As we made used of adaptivity, the corresponding moment orders and used quadrature points can best be seen in Table \ref{tab:orders3D}.
\begin{table}
	\begin{tabular}{l|cccc}
		Sparse grids&&&&\\\hline
		\quad Moment order&1&2&-&-\\
		\quad Number of quadrature points&25&441&-&-\\
		Tensorized grids & & & & \\\hline
		\quad Moment order&1&2&3&4\\
		\quad Number of quadrature points&8&27&64&125\\
	\end{tabular}
	\caption{Order of moments and corresponding number of quadrature nodes used.}
	\label{tab:orders3D}
\end{table}
The following properties emerge:
\begin{itemize}
	\item It was not possible to obtain a result for sparse grids with moments of order 3 or higher. Any level of the used sparse grids based on Clenshaw Curtis quadrature tested up to level 11 ($72\;705$ quadrature nodes) for order 3, resulted in a failure of the method while computing the Hessian during the Newton iteration.
	\item Sparse grids required a significant higher number of quadrature points for the sample order of moment in comparison to tensorized grids.
	\item When using sparse grids with adaptivity the computational cost per quadrature node was observed to be smaller, proving that the nestedness property can potentially increase the methods performance.
\end{itemize}
Even though the solutions in Figure \ref{fig:solution3D} show the same characteristics and sparse grids are generally well suited and often used in combination with the SC method, our tests reveal that, at least for the Clenshaw-Curtis based sparse grids we have used, these quadrature rules are not well suited to be used in combination with IPM as they require significantly more quadrature points and only give results for low orders of moments. However, as three stochastical dimensions are difficult to deal with in hyperbolical problems, but not for an UQ analysis in general, this might change for high dimensions.
\begin{figure}[H]
		\includegraphics[scale=0.18]{figs/pipe/pipe_sc_ref_n50_E(rho).png}
		\label{fig:referenceSolutionsPipe1}
		\includegraphics[scale=0.18]{figs/pipe/pipe_sc_ref_n50_Var(rho).png}
		\label{fig:referenceSolutionsPipe2}
		\includegraphics[scale=0.18]{figs/pipe/pipe_mesh.png}
		\label{fig:referenceSolutionsPipeMesh}
		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_ERho.png}
		\label{fig:adIPMSolutionsPipeSGE}
		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_VarRho.png}
		\label{fig:adIPMSolutionsPipeSGVar}
		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_adipm_n1-2_refinementLevel.png}
		\label{fig:adIPMSolutionsPipeSGRI}
		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_ERho.png}
		\label{fig:adIPMSolutionsPipeTGE}
		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_VarRho.png}
		\label{fig:adIPMSolutionsPipeTGVar}
		\includegraphics[scale=0.18]{figs/pipe/euler2D_pipe_n1-4_sg4-6_refinementLevel.png}
		\label{fig:adIPMSolutionsPipeTGRI}
	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh (top row) and corresponding adaptive IPM solutions for sparse (middle row) and tensorized grids (bottom row) with the used refinement levels at the last time step}
	\label{fig:solution3D}
